My wonderful paper

### 2024-11-07 _manuscript/_tex/index.tex
add more dot points

@@

In this paper, we formalize these informal checks into an \emph{analysis
plan} comprising the analysis steps along with the associated unit
tests. This formalization allows us to examine assumptions made about
the data during analysis and to compare different unit tests in
diagnosing unexpected outcomes. We then introduce a procedure to assess
the quality of these unit tests based on their accuracy and redundancy
across simulated version of the original data. The accuracy is assessed
using binary classification metrics -- precision and recall -- while
redundancy is measured using mutual information. We illustrate these
concepts with a simple step count example and a scenario of diagnosing
unexpected regression results when studying the effect of PM10 on
mortality.

@@

Currently, little has been done on how to computationally incorporate
this diagnostic approach into data analysis workflow or software. Most
of the diagnostic tools focus on defining user-specified rules, such as
data quality checks or producing metrics to summarize model performance,
as in model diagnostics. For example, the \texttt{assertr} package
\citep{assertr} and the \texttt{pointblank} package \citep{pointblank}
provide data validation tools to allow users to define data quality
checks. In contrast, packages provide model checks tools like
\texttt{performance} \citep{performance} and \texttt{yardstick}
\citep{yardstick}, from the \texttt{tidymodels} \citep{tidymodels}
ecosystem, offers goodness-of-fit, residual analysis, and performance
metrics.

@@

Software testing relies on ``oracles'' to provide the expected output
necessary for verifying test results. For example, to test whether the
program correctly calculates 1 + 1, one need to supply the correct
answer, 2. However, establishing this ``correct'' output can sometimes
be challenging, where obtaining a solution may be difficult without the
program itself. This situation leads to the oracle problem
\citep{barr2014oracle}. In data analysis, the similar oracle problem can
happen, as the ``truth'' of an outcome, the expectation, depends on the
underlying theory or interpretation. For example, in a linear regression
model, the significance of a coefficient may be expected or unexpected
based on the theory, making it challenging for researcher with a
different theory to assess the results and the analysis.

@@

\begin{itemize}
\item
  how to systematically simulate data is still unknown, sensitivity of
  the simulation to the results
\item
  plotting is a critical way to check data and they can still be frame
  into a unit test. it is a open problem to how to encode the
  visualization into the unit tests. Maybe a procedure like confirm plot
  (this looks alright to you) and then press the button to continue
\item
  currently no automated way to generate unit tests. It is interesting
  to see the automation of generating unit tests, although it requires
  the inputs from experts across a wide array of common scenarios.
\end{itemize}

>>>>

\begin{itemize}
\tightlist
\item
  emphasize trustworthy data science since it is the theme of the
  special issue
\end{itemize}

@@

In this paper, we formalize these informal checks into an \emph{analysis
plan} comprising the analysis steps along with the associated unit
tests. This formalization allows us to examine assumptions made about
the data during analysis and to compare different unit tests in
diagnosing unexpected outcomes. We then introduce a procedure to assess
the quality of these unit tests based on their accuracy and redundancy
across simulated version of the original data. The accuracy is assessed
using binary classification metrics -- precision and recall -- while
redundancy is measured using mutual information. The workflow provides
numerical guarantee of the consistency of the expectation and the data
for the analysis plan. This means that given the assumption of the data
generating mechanism, the analysis will provide the outcome we expect.
In data analysis, analysts are not necessarily collecting the data and
they can only make reasonable assumption about the quality of the data.
This method we propose provides a numerical guarantee, based on
simulation, that the analysis will produce the outcome we expect, given
the assumption of the data generating mechanism. This provides
transparency of the analysis and build trustworthy data analysis.

@@

Currently, little has been done on how to computationally incorporate
this diagnostic approach into data analysis workflow or software. Most
of the diagnostic tools focus on defining user-specified rules, such as
data quality checks or producing metrics to summarize model performance,
as in model diagnostics. For example, the \texttt{assertr} package
\citep{assertr} and the \texttt{pointblank} package \citep{pointblank}
provide data validation tools to allow users to define data quality
checks. In contrast, packages provide model checks tools like
\texttt{performance} \citep{performance} and \texttt{yardstick}
\citep{yardstick}, from the \texttt{tidymodels} \citep{tidymodels}
ecosystem, offers goodness-of-fit, residual analysis, and model
performance metrics.

@@

@@

\begin{itemize}
\item
  how to systematically simulate data is still unknown, sensitivity of
  the simulation to the results
\item
  plotting is a critical way to check data and they can still be frame
  into a unit test. it is a open problem to how to encode the
  visualization into the unit tests. Maybe a procedure like confirm plot
  (this looks alright to you) and then press the button to continue
\item
  currently no automated way to generate unit tests. It is interesting
  to see the automation of generating unit tests, although it requires
  the inputs from experts across a wide array of common scenarios.
\item
  There are cost and benefit on what granularity to set expectation.
  Setting them on every step of the analysis requires more work from the
  analysts while for complex analysis, the diagnostics may require
  divide the analysis into different components and diagnose it in a
  hierarchical structure. The method proposed in this paper should be
  applicable for both fine and coarse analysis.
\item
  (only mention it) Software testing relies on ``oracles'' to provide
  the expected output necessary for verifying test results. For example,
  to test whether the program correctly calculates 1 + 1, one need to
  supply the correct answer, 2. However, establishing this ``correct''
  output can sometimes be challenging, where obtaining a solution may be
  difficult without the program itself. This situation leads to the
  oracle problem \citep{barr2014oracle}. In data analysis, the similar
  oracle problem can happen, as the ``truth'' of an outcome, the
  expectation, depends on the underlying theory or interpretation. For
  example, in a linear regression model, the significance of a
  coefficient may be expected or unexpected based on the theory, making
  it challenging for researcher with a different theory to assess the
  results and the analysis.
\end{itemize}

### 2024-11-07 index.qmd
add more dot points

@@

In this paper, we formalize these informal checks into an *analysis plan* comprising the analysis steps along with the associated unit tests. This formalization allows us to examine assumptions made about the data during analysis and to compare different unit tests in diagnosing unexpected outcomes. We then introduce a procedure to assess the quality of these unit tests based on their accuracy and redundancy across simulated version of the original data. The accuracy is assessed using binary classification metrics -- precision and recall -- while redundancy is measured using mutual information. We illustrate these concepts with a simple step count example and a scenario of diagnosing unexpected regression results when studying the effect of PM10 on mortality.

@@

Software testing relies on "oracles" to provide the expected output necessary for verifying test results. For example, to test whether the program correctly calculates 1 + 1, one need to supply the correct answer, 2. However, establishing this "correct" output can sometimes be challenging, where obtaining a solution may be difficult without the program itself. This situation leads to the oracle problem [@barr2014oracle]. In data analysis, the similar oracle problem can happen, as the "truth" of an outcome, the expectation, depends on the underlying theory or interpretation. For example, in a linear regression model, the significance of a coefficient may be expected or unexpected based on the theory, making it challenging for researcher with a different theory to assess the results and the analysis.

@@

>>>>

- emphasize trustworthy data science since it is the theme of the special issue

@@

In this paper, we formalize these informal checks into an *analysis plan* comprising the analysis steps along with the associated unit tests. This formalization allows us to examine assumptions made about the data during analysis and to compare different unit tests in diagnosing unexpected outcomes. We then introduce a procedure to assess the quality of these unit tests based on their accuracy and redundancy across simulated version of the original data. The accuracy is assessed using binary classification metrics -- precision and recall -- while redundancy is measured using mutual information. The workflow provides numerical guarantee of the consistency of the expectation and the data for the analysis plan. This means that given the assumption of the data generating mechanism, the analysis will provide the outcome we expect. In data analysis, analysts are not necessarily collecting the data and they can only make reasonable assumption about the quality of the data. This method we propose provides a numerical guarantee, based on simulation, that the analysis will produce the outcome we expect, given the assumption of the data generating mechanism. This provides transparency of the analysis and build trustworthy data analysis.

@@

@@

* There are cost and benefit on what granularity to set expectation. Setting them on every step of the analysis requires more work from the analysts while for complex analysis, the diagnostics may require divide the analysis into different components and diagnose it in a hierarchical structure. The method proposed in this paper should be applicable for both fine and coarse analysis.

* (only mention it) Software testing relies on "oracles" to provide the expected output necessary for verifying test results. For example, to test whether the program correctly calculates 1 + 1, one need to supply the correct answer, 2. However, establishing this "correct" output can sometimes be challenging, where obtaining a solution may be difficult without the program itself. This situation leads to the oracle problem [@barr2014oracle]. In data analysis, the similar oracle problem can happen, as the "truth" of an outcome, the expectation, depends on the underlying theory or interpretation. For example, in a linear regression model, the significance of a coefficient may be expected or unexpected based on the theory, making it challenging for researcher with a different theory to assess the results and the analysis.
